Galoissche Kennzeichnung von Funktionenkörper \\ vom Transzendenzgrad $1$ über Zahlkörper

### 2017-05-20 Master Seminar/article.tex
Add master seminar tex file

\chapter{Valuations}

\section{The Valuation Spectra}

In this section we shall recall facts from valuation theory.

@@

\begin{remark}
\begin{itemize}
\item A one-dimensional finitely generated field is just a global field. 
\item A two-dimensional finitely generated field over $\mathbb{Q}$ is a function field in one variable over $\mathbb{Q}$.
\end{itemize}
\end{remark}

@@

We denote the valuation ring of $v\in\operatorname{Spv}(K)$ by $(\mathcal{O}_v,\mathfrak{m}_v)$ and set $U_v=\mathcal{O}_v^\times$ and let $k(v)=\mathcal{O}_v/\mathfrak{m}_v$ be its residue field. The \textit{rank} of $v$ is defined as the Krull dimension of its valuation ring $\mathcal{O}_v$. The \textit{rational rank}\index{valuation!rational rank} of $v:K\to\Gamma_v\cup\{\infty\}$ is defined as the rank of its value group $\Gamma_v$ as a $\mathbb{Z}$-module:
\[\operatorname{rrank}(v) = \dim_\mathbb{Q}(\Gamma_v\otimes_\mathbb{Z}\mathbb{Q}) \]
\end{definition}

@@

\begin{proof}
See \cite{EP05} 3.4.1 and 3.4.3.
\end{proof}

@@

\begin{definition}
We can endow $\operatorname{Spv}(K)$ with a natural topology by declaring following sets as a basis of open sets:
\[\{v\in\operatorname{Spv}(K)\mid v(x)\geq 0\text{ for all }x\in S\} \]
where $S$ ranges over all finite subsets of $K$. 
\end{definition}

\begin{lemma}\label{spv-topology}
$\operatorname{Spv}(K)$ with the topology defined above is a quasi-compact $T_0$-space. In particular the set of closed points $|\operatorname{Spv}(K)|$ of $\operatorname{Spv}(K)$ is a quasi-compact $T_1$-space, endowed with the induced topology. Let $v,w\in\operatorname{Spv}(K)$. We have:
\[\mathcal{O}_w\subset\mathcal{O}_v \iff w\in\overline{\{v\}} \]
Thus $v$ is a closed point if and only if it has no non-trivial specializations.
\end{lemma}

\begin{proof}
See \cite{ZS60} VI §17 Theorem 38, §17 Theorem 40.
\end{proof}

\begin{lemma}\label{2.2-pre}
For a fixed $v\in \operatorname{Spv}(K)$ there is a canonical bijection: 
\begin{align*}
\{ w\in\operatorname{Spv}(K)\mid \mathcal{O}_w\subset\mathcal{O}_v \} &\longrightarrow \operatorname{Spv}k(v)\\ 
w&\longmapsto w\mod v
\end{align*}
Therefore $v$ is a closed point iff $k(v)$ only admits the trivial valuation, i.e. $k(v)$ has characteristic $p>0$ and is algebraic over $\mathbb{F}_p$. In other words, $v$ is a closed point iff $\dim k(v)=0$.

If $w\in\operatorname{Spv}_\text{rd}^{(i)}(K)$ for some $i\geq 1$ then there exists a unique $v(w)\in\operatorname{Spv}_\text{rd}^{(i-1)}(K)$ such that $\mathcal{O}_w\subset\mathcal{O}_{v(w)}$. $w$ is henselian if and only if $v(w)$ and $w\mod v(w)$ are henselian.
\end{lemma}

\begin{proof}
See \cite{EP05} 2.3, 4.1.4, \cite{ZS60} VI §3 Theorem 3.
\end{proof}

\begin{theorem}\label{theorem-henselization}
Let $K$ be a field and $v\in\operatorname{Spv}(K)$. Then there is a separable extension $K_v/K$ with a henselian extension $\overline{v}$ of $v$ with the following characterization:

\textit{If $K'/K$ is henselian with respect to an extension $w$ of $v$, then there exists a uniquely determined $K$-embedding $K_v\to K'$, so that $w$ extends $\overline{v}$.}

$K_v$ is called a \textit{henselization} of $K$ with respect to $v$. Furthermore, $\overline{v}$ has the same value group as $v$ and the same residue field $k(\overline{v}) = k(v)$. $G_{K_v}$ is the decomposition group of the unique extension of $\overline{v}$ to $K^\text{s}$:
\[ G_{\overline{v}} = \{ \sigma\in G_K\mid \sigma(\mathcal{O}_{\overline{v}}) =\mathcal{O}_{\overline{v}} \} \subset G_K \]
Conversely, taking the fixed field of the decomposition group of an extension $\overline{v}'$ of $v$ in $K^\text{s}$ yields a henselization of $K$ with respect to $v$. Any two of such henselizations are $K$-conjugate.
\end{theorem}

\begin{proof}
See \cite{EP05} Lemma 5.2.1, Lemma 5.2.2, Theorem 5.2.5.
\end{proof}

\begin{theorem}\label{theorem-absoluteinertia}
Let $K$ be a henselian field with respect to $v: K\to\Gamma\cup\{\infty\}$ with unique extension $\overline{v}$ on $K^\text{s}$. Let $K^\text{t}$ be the absolute inertia field of $K$ with the unique extension $w$ of $v$, i.e. $K^\text{t}$ is the fixed field of the absolute inertia group $T$: 
\[ T = \{ \sigma\in G_K\mid \sigma(x)\equiv x\mod \mathfrak{m}_{\overline{v}} \quad \text{for all }x\in\mathcal{O}_{\overline{v}} \}\lhd G_K \]
Then we have:
\begin{enumerate}[(i)]
\item $w$ has value group $\Gamma$ and residue class field $k(v)^\text{s}$.
\item $K^\text{t}$ is a Galois extension of $K$ and $\operatorname{Gal}(K^\text{t}/K)\cong G_{k(v)}$ as profinite groups. There exists an inclusion-preserving one-to-one correspondence: 
\begin{align*}
\{ \text{Intermediate extensions in $K^\text{t}/K$} \} &\longrightarrow \{\text{Separable extensions of $k(v)$} \}\\ 
L &\longmapsto k(v_L)  
\end{align*}
where $v_L$ is the unique extension of $v$ to $L$.
\end{enumerate}
\end{theorem}

\begin{proof}
See \cite{EP05} Theorem 5.2.7.
\end{proof}

\begin{lemma}\label{lemma-henselization-specialization}
Let $K$ be a field and $w\in\operatorname{Spv}(K)$ with a fixed henselization $K_w$ of $w$. If $v\in\operatorname{Spv}(K)$ such that $w$ is a specialization of $v$, then there is a henselization $K_v$ of $v$ such that $K_v\subset K_w\subset K_v^\text{t}$.
\end{lemma}

\begin{proof}
Let $\overline{w}$ be the extension of $w\in\operatorname{Spv}(K_w)$ on $K^\text{s}$. By \cite{EP05} Lemma 3.1.5 we can find an extension $\overline{v}$ of $v$ on $K^\text{s}$ such that $\overline{w}$ is a specialization of $\overline{v}$. Let $K_v$ be the fixed field of $G_{\overline{v}}\subset G_K$.

Let $\sigma\in G_{\overline{w}}$. Then both $\mathcal{O}_{\overline{v}}$ and $\sigma(\mathcal{O}_{\overline{v}})$ are overrings of $\mathcal{O}_{\overline{w}}$. Now the overrings of valuation rings are totally ordered by inclusion and correspond to localizations of $\mathcal{O}_{\overline{w}}$ at prime ideals (See \cite{EP05} 2.3). Since $\mathcal{O}_{\overline{v}}$ and $\sigma(\mathcal{O}_{\overline{v}})$ have the same Krull dimension, they must be the same, thus $\sigma\in G_{\overline{v}}$. This proves $K_v\subset K_w$.

Now let $\sigma\in G_{K_v^\text{t}}$, i.e. $\sigma\in G_{\overline{v}}$ such that the induced automorphism on the residue field $\overline{\sigma}:k(\overline{v})\to k(\overline{v})$ is the identity. Let $\mathfrak{p}\subset\mathcal{O}_{\overline{w}}$ be a prime ideal with $\mathcal{O}_{\overline{v}} = (\mathcal{O}_{\overline{w}})_\mathfrak{p}$. In particular, the restriction $\overline{\sigma}|: \mathcal{O}_{\overline{w}}/\mathfrak{p}\to \mathcal{O}_{\overline{w}}/\mathfrak{p}$ is the identity, thus $\sigma(\mathcal{O}_{\overline{w}})=\mathcal{O}_{\overline{w}}$, i.e. $\sigma\in G_{\overline{w}}$. This proves $K_w\subset K_v^\text{t}$.
\end{proof}

\begin{theorem}\label{theorem-sylowgroups-inertiagroup}
Let $K$ be a henselian field with respect to $v:K\to\Gamma\cup\{\infty\}$ with unique extension $\overline{v}$ on $K^\text{s}$, and $p=\operatorname{char}k(v)$. Then the absolute ramification group
\[ V = \Big\{ \sigma\in G_K\mathop{\Big|} \frac{\sigma(x)}{x}-1\in\mathfrak{m}_{\overline{v}}\quad \text{for all }x\in {K^\text{s}}^\times \Big\} \lhd T \]
is the unique $p$-Sylow group of the absolute inertia group $T$. $V$ is trivial if $p=0$. We have the following isomorphism of profinite groups:
\[ T/V \cong \prod_{\ell\neq p\text{ prime}}\mathbb{Z}_\ell^{r_\ell} \]
where $r_\ell = \dim_{\mathbb{F}_\ell}(\Gamma/\ell\Gamma)$.
\end{theorem}

\begin{proof}
See \cite{EP05} Theorem 5.3.3.
\end{proof}

\begin{theorem}[\textit{F.K. Schmidt}]\label{theorem-fkschmidt}
Let $v_1,v_2\in\operatorname{Spv}(K)$ be two independent henselian valuations on a field $K$, i.e. $\mathcal{O}_{v_1}\mathcal{O}_{v_2}=K$. Then $K$ is separably closed.
\end{theorem}

\begin{proof}
See \cite{EP05} Theorem 4.4.1.
\end{proof}

\begin{remark}
Two valuations of rank $1$ are independent iff they are inequivalent.
\end{remark}

\begin{corollary}\label{corollary-fkschmidt}
Let $w_1,w_2\in\operatorname{Spv}(K^\text{s})$ be two inequivalent valuations of rank $1$ with decomposition groups $G_{{w_1}},G_{{w_2}}\subset G_K$. Then $G_{{w_1}}\cap G_{{w_2}}=1$.
\end{corollary}

\begin{proof}
Let $L$ be the fixed field of $G_{w_1}\cap G_{w_2}\subset G_K$ and set $v_i = w_i|_L$. Then $L$ is henselian with respect to both $v_1$ and $v_2$. Since $w_1,w_2$ are the only extensions of $v_1$ and $v_2$ in $K^\text{s}$ respectively, $v_1$ and $v_2$ are inequivalent valuations of rank $1$. Now $L$ is separably closed by Theorem \ref{theorem-fkschmidt}. Therefore, by Theorem \ref{theorem-henselization}, we get $G_{{w_1}}\cap G_{{w_1}}=G_L =1$.
\end{proof}

\section{The Main Lemma}

Let $K$ be a (not necessarily finitely generated) field over $\mathbb{Q}$ and of dimension $2$.

\begin{lemma}\phantomsection\label{lemma-defectless-valuations}
\begin{enumerate}[(i)]
\item $\operatorname{Spv}_\text{d}^{(1)}(K) = \operatorname{Spv}_\text{rd}^{(1)}(K)$
\item $\operatorname{Spv}^{(2)}(K) = \operatorname{Spv}_\text{d}^{(2)}(K)\subset |\operatorname{Spv}(K)|$
\end{enumerate}
\end{lemma}

\begin{proof}
\begin{enumerate}[(i)]
\item Let $v$ be a defectless valuation of rational rank $1$. By Lemma \ref{rank-dimension-inequality}, $\operatorname{rank}(v) \leq \operatorname{rrank}(v) = 1$. Since $\Gamma_v$ is not trivial, $\operatorname{rank}(v) = 1=\operatorname{rrank}(v)$.
\item Let $v$ be a valuation of rational rank $2$. Due to Lemma \ref{rank-dimension-inequality}, the dimension inequality is an equality and we have $\dim k(v) = 0$. \qedhere
\end{enumerate}
\end{proof}

\begin{lemma}\label{2.2}
Let $K$ be henselian with respect to two different valuations $w_1,w_2\in |\operatorname{Spv}(K)|$. Then $K$ is algebraically closed or $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K),\ v(w_1)=v(w_2)$ and $k(v(w_1))$ is separably closed. In this case, we have $\operatorname{cd}(G_K)\leq 2$.
\end{lemma}

\begin{proof}
Assume that $K$ is not algebraically closed. $w_1$ and $w_2$ can't be trivial, for otherwise $\dim k(w_i)=\dim (K)\neq 0$. By the Theorem of F.K. Schmidt \ref{theorem-fkschmidt}, we can see that $w_1,w_2$ are dependent. Since two different valuations of rank $1$ are independent, at least one of $w_1$ and $w_2$ is of rank $2$. Now, assume that $w_1$ is, say, of rank $1$ and $w_2$ is of rank $2$. Since they are dependent, it follows $v(w_2) = w_1$ and $w_1$ has a non-trivial specialization, a contradiction to $w_1\in |\operatorname{Spv}(K)|$.

Thus $w_1,w_2$ are both of rank $2$ and by Lemma \ref{rank-dimension-inequality} we can see that $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$. Set $v_i=v(w_i)\in\operatorname{Spv}^{(1)}_\text{d}(K)$. By Lemma \ref{2.2-pre} $K$ is henselian with respect to both $v_1$ and $v_2$. Again, by the Theorem of F.K. Schmidt \ref{theorem-fkschmidt}, we get $v_1 = v_2$. Furthermore $k(v_1) = k(v_2)$ is henselian with respect to two different valuations of rank $1$, namely $w_1\mod v_1$ and $w_2\mod v_1$, making $k(v_1)$ separably closed.

The second assertion follows from the first and from [K2], §5, Theorem 1.
\end{proof}

\begin{definition}
Let $K$ be finitely generated. A \textit{model} of $K$ is a regular, proper scheme $\mathfrak{X}$ over $\mathbb{Z}$ with function field $K$. $\mathfrak{X}_i$ denotes the set of points $x\in\mathfrak{X}$ whose closure $\overline{\{x\}}$ is $i$-dimensional.

A \textit{morphism of models} is a morphism of schemes inducing the identity on $K$. We denote the category of models of $K$ by $\mathfrak{M}(K)$; it is filtered.

For every model $\mathfrak{X}$ there is a canonical map $\pi_\mathfrak{X}:\operatorname{Spv}(K)\to\mathfrak{X}$. $\pi_\mathfrak{X}(v)$  is the unique point of $\mathfrak{X}$ such that $\mathcal{O}_{\mathfrak{X}, \pi_\mathfrak{X}(v)}\subset\mathcal{O}_v$ and the inclusion $\mathcal{O}_{\mathfrak{X}, \pi_\mathfrak{X}(v)}\hookrightarrow \mathcal{O}_v$ is a homomorphism of local rings.
\end{definition}

\begin{lemma}\phantomsection\label{2.3}
\begin{enumerate}[(i)]
\item $\operatorname{Spv}(K)$, together with the maps $\pi_{\mathfrak{X}}$, is the following projective limit in the category of topological spaces: 
\[\operatorname{Spv}(K) =\varprojlim_{\mathfrak{X}\in\mathfrak{M}(K)}\mathfrak{X}\]
In particular we have:
\[ |\operatorname{Spv}(K)| =\varprojlim_{\mathfrak{X}\in\mathfrak{M}(K)}|\mathfrak{X}| \]
\item For all $v\in\operatorname{Spv}(K)$ we have:
\[ \mathcal{O}_v = \varinjlim_{\mathfrak{X}\in\mathfrak{M}(K)}\mathcal{O}_{\mathfrak{X},\pi_\mathfrak{X}(v)} \]
\item For $\mathfrak{X}\in\mathfrak{M}(K),\ x\in\mathfrak{X}$ let $K_x$ be the quotient field of a henselization of $\mathcal{O}_{\mathfrak{X},x}$. Then we have:
\[ K_v=\varinjlim_{\mathfrak{X}\in\mathfrak{M}(K)}K_{\pi_\mathfrak{X}(v)} \]
\end{enumerate}
\end{lemma}

\begin{proof}
(i) and (ii) follow from [HK], Lemma 2.4.3 and (iii) is a consequence of (ii) and the fact that henselizations commute with filtrated direct limits, see \cite{EGA4} 18.6.14.
\end{proof}

The following lemma is the key result in proving Theorem \ref{thm:main-result}. First, we use K. Kato's two-dimensional local-global principle to show the assertion for finitely generated $K$. Then, we'll use a quasi-compactness argument to see that we can drop the assumption that $K$ is finitely generated.

\begin{lemma}\label{2.4-finitelygenerated}
Let $K$ be finitely generated and $\ell\neq 2$ be a prime. The following map, induced by the restrictions, is injective:
\[ \mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z})\longrightarrow\prod_{w\in |\operatorname{Spv}(K)| }\mathrm{H}^3(K_w,\mathbb{Z}/\ell\mathbb{Z}) \]
\end{lemma}

\begin{proof}
Let $L=K(\mu_{\ell},\sqrt{-1})/K$. Consider the restriction and corestriction maps:
\[ \begin{tikzcd}
\mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z}) \ar[r, "\operatorname{res}"] & \mathrm{H}^3(L,\mathbb{Z}/\ell\mathbb{Z})\ar[r, "\operatorname{cor}"]& \mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z}) 
\end{tikzcd} \]
We have $\operatorname{cor}\circ\operatorname{res} = [L:K]$. Since $[L:K]$ is coprime to $\ell$, we see that $\operatorname{res}$ is injective. Thus we can assume that $K=L$, i.e. $\mu_{\ell}\subset K$ and $K$ is not formally real. Now we can replace $\mathbb{Z}/\ell\mathbb{Z}$ by $\mathbb{Z}/\ell\mathbb{Z}(2) = \mu_\ell^{\otimes 2}$ and let us write $\mathrm{H}^3(L) = \mathrm{H}^3(L, \mathbb{Z}/\ell\mathbb{Z}(2))$.

Let $\mathfrak{X}$ be a model of $K$. For $x\in\mathfrak{X}_1$ the local ring $\mathcal{O}_{\mathfrak{X}, x}$ is a discrete valuation ring and defines a valuation $v_x\in\operatorname{Spv}_\text{d}^{(1)}(K)$. By \cite{Ka86} Theorem 0.8, the map
\[ \mathrm{H}^3(K)\longrightarrow \prod_{x\in\mathfrak{X}_1}\mathrm{H}^3(K_{v_x}) \]
is injective and its image is contained in $\bigoplus \mathrm{H}^3(K_{v_x})$. By \cite{Ka86} Lemma 1.4, for a discrete valuation $v$ of $K$ the map
\[ \mathrm{H}^3(K_v) \longrightarrow \prod_w\mathrm{H}^3(K_w) \]
where $w$ ranges over all rank-$2$ valuations with $v(w) = v$, can be identified with the following map:
\[ {}_\ell\operatorname{Br}(k) \longrightarrow \prod_{\widetilde{w}\in\operatorname{Spv}^{(1)}(k)} {}_\ell\operatorname{Br}(k_{\widetilde{w}}) \]
where $k = k(v)$. This map is injective by the classical Hasse principle and its image is contained in $\bigoplus {}_\ell\operatorname{Br}(k_{\widetilde{w}})$. Combining these maps yields the injection:
\[\Phi: \mathrm{H}^3(K)\longrightarrow \bigoplus_{x\in\mathfrak{X}_1}\  \bigoplus_{\substack{w\in\operatorname{Spv}_\text{rd}^{(2)}(K) \\ v(w) = v_x}} \mathrm{H}^3(K_w) \]
The direct sum ranges over a subset of $|\operatorname{Spv}(K)|$, thus this concludes the proof. 
\end{proof}

In order to generalize Lemma \ref{2.4-finitelygenerated} for arbitrary, not necessarily finitely generated $K$, we need the following Lemma:

\begin{lemma}\label{2.4-closedset}
Let $f\in\mathrm{H}^3(K)$. Then the support of $f$ is closed in $|\operatorname{Spv}(K)|$:
\[ \operatorname{supp}(f) = \{w\in |\operatorname{Spv}(K)|\mid \operatorname{res}_{K_w/K}(f)\neq 0 \} \]
where $\operatorname{res}_{K_w/K}: \mathrm{H}^3(K)\to \mathrm{H}^3(K_w)$.
\end{lemma}

\begin{proof}
We adopt the notation from the previous proof. First, we shall prove that the following map is injective for every model $\mathfrak{X}$ of $K$:
\[ f: \mathrm{H}^3(K) \longrightarrow \prod_{y\in\mathfrak{X}_0}\mathrm{H}^3(K_y) \]
For $x\in\mathfrak{X}$ let $\mathcal{O}_x$ be a henselization of $\mathcal{O}_{\mathfrak{X}, x}$ and $K_x$ its quotient field. For $y\in\mathfrak{X}_0$ set $P_y=\operatorname{Spec}(\mathcal{O}_y)_1$. For $z\in P_y$ let $\mathcal{O}_z$ be a henselization of $\mathcal{O}_y$ at $z$ and $K_z$ its quotient field. For $y\in\mathfrak{X}_0$ and $x\in\mathfrak{X}_1$ let $P_{x,y}$ be the subset of $P_y$ consisting of all $z\in P_y$ which are mapped to $x$ under $\operatorname{Spec}(\mathcal{O}_y)\to\mathfrak{X}$. Consider the following commutative diagram:
\[ \begin{tikzcd}
\mathrm{H}^3(K)\ar[r, "\varphi"]\ar[d, "f"'] & \displaystyle\prod_{x\in\mathfrak{X}_1}\mathrm{H}^3(K_x)\ar[d, "\psi"]\\
\displaystyle\prod_{y\in\mathfrak{X}_0}\mathrm{H}^3(K_y) \ar[r, "g"']& \displaystyle\prod_{\substack{x\in\mathfrak{X}_1 \\ y\in\mathfrak{X}_0}}\prod_{z\in P_{x,y}}\mathrm{H}^3(K_z)
\end{tikzcd} \]
According to [Sa], Theorem (2-15) $g$ is injective. Now, $\psi\circ\varphi$ can be identified with the injective map $\Phi$ from the proof of the previous Lemma. Thus its image is contained in $\bigoplus\mathrm{H}^3(K_w)$. Therefore, $f$ is injective as well and its image is contained in $\bigoplus\mathrm{H}^3(K_y)$. Now, one easily checks that:
\[ \operatorname{supp}(f) = \bigcap_{\mathfrak{X}\in \mathfrak{M}(K)} \pi_\mathfrak{X}^{-1}(\{ y\in\mathfrak{X}_0\mid \operatorname{res}_{K_y/K}(f)\neq 0 \}) \]
Note that the right set is closed in $\operatorname{Spv}(K)$.
\end{proof}

Now the general case follows by a quasi-compactness argument:

\begin{lemma}[\textit{Main Lemma}]\label{2.4}
Let $K$ be not necessarily finitely generated and $\ell\neq 2$ be a prime. The following map, induced by the restrictions, is injective:
\[ \mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z})\longrightarrow \prod_{w\in |\operatorname{Spv}(K)| }\mathrm{H}^3(K_w,\mathbb{Z}/\ell\mathbb{Z}) \]
\end{lemma}

\begin{proof}
Just like in Lemma \ref{2.4-finitelygenerated} we can assume that $\mu_\ell\subset K$ and $K$ is not formally real and replace $\mathbb{Z}/\ell\mathbb{Z}$ by $\mathbb{Z}/\ell\mathbb{Z}(2)$. We adopt the notation from the previous proofs.

Let $f\in\mathrm{H}^3(K)$ such that $\operatorname{res}_{K_w/K}(f) = 0$ for all $w\in |\operatorname{Spv}(K)|$. There exists a finitely generated subfield $K_0$ of $K$ such that $f = \operatorname{res}_{K/K_0}(g)$ for some $g\in\mathrm{H}^3(K_0)$ and $K/K_0$ is algebraic. Let $\{K_i\mid i\in I\}$ be the set of all finite subextensions of $K/K_0$. For $w\in |\operatorname{Spv}(K)|$ we denote the restriction of $w$ to $K_i$ with $w$ as well.

Let  $w\in |\operatorname{Spv}(K)|$. Since $\mathrm{H}^3(K_w) = \varinjlim_I \mathrm{H}^3((K_i)_w)$ there exists an $i = i(w)\in I$ such that $\operatorname{res}_{(K_i)_w/K}(g) = 0$. By Lemma \ref{2.4-closedset} there is an open neighborhood $U_{i(w)}\subset |\operatorname{Spv}(K_i)|$ of $w$ such that $\operatorname{res}_{(K_i)_v/K}(g)=0$ for all $v\in U_{i(w)}$. For $K_i\subset K_j\subset K$ let
\[ \varphi_i: |\operatorname{Spv}(K)| \to |\operatorname{Spv}(K_i)| \]
\[ \varphi_{ij}: |\operatorname{Spv}(K_j)|\to |\operatorname{Spv}(K_i)| \]
be the restriction maps. They are continuous. Since $|\operatorname{Spv}(K)|$ is quasi-compact by Lemma \ref{spv-topology} we can find a finite set $S\subset |\operatorname{Spv}(K)|$ such that:
\[ \bigcup_{w\in S}\varphi^{-1}_{i(w)}(U_{i(w)}) = |\operatorname{Spv}(K)| \]
Let $j\in I$ such that $K_j$ is the composite of the fields $K_{i(w)},\ w\in S$. We also have:
\[ \bigcup_{w\in S}\varphi^{-1}_{i(w)j}(U_{i(w)}) = |\operatorname{Spv}(K_j)| \]
Let $v\in |\operatorname{Spv}(K_j)|$ and let $i = i(w),\ w\in S$ such that $v\in\varphi_{ij}^{-1}(U_{i(w)})$. We have:
\[ \operatorname{res}_{(K_j)_v/K_j}(\operatorname{res}_{K_j/K_0}(g)) = \operatorname{res}_{(K_j)_v/(K_i)_v}(\operatorname{res}_{(K_i)_v/K_0}(g)) = 0 \]
Since $K_j$ is finitely generated we obtain $\operatorname{res}_{K_j/K_0}(g) = 0$ with Lemma \ref{2.4-finitelygenerated}, thus $f = \operatorname{res}_{K/K_0}(g) = 0$.
\end{proof}

Now we calculate the cohomology groups $\mathrm{H}^3(K_w,\mathbb{Z}/\ell\mathbb{Z})$ for $w\in |\operatorname{Spv}(K)|$. With the help of Lemma \ref{2.4} we can find a cohomological criterion of henselian fields over $K$ with respect to a $w\in\operatorname{Spv}_\text{d}^{(2)}(K)$.

\begin{lemma}\label{2.5}
Let $K$ be henselian with respect to $w\in|\operatorname{Spv}(K)|$. Set $k=k(w)$ and $\Gamma = \Gamma_w$. Then $k$ is an algebraic extension of a finite field $\mathbb{F}_p$ by Lemma \ref{2.2-pre}. Let $\ell\neq p$ be a prime and assume $\mu_\ell\subset K$. Then:
\[ \mathrm{H}^3(K, \mathbb{Z}/\ell\mathbb{Z}) = \begin{cases}
\mathbb{Z}/\ell\mathbb{Z}, & \ell^\infty\nmid [k:\mathbb{F}_p] \text{ and } \dim_{\mathbb{F}_\ell}(\Gamma / \ell\Gamma) = 2\\
0, & \text{otherwise}
\end{cases} \]
If $L/K$ is an algebraic extension and $\mathrm{H}^3(K, \mathbb{Z}/\ell\mathbb{Z}) = \mathbb{Z}/\ell\mathbb{Z}$ holds, we have:
\[ \mathrm{H}^3(L,\mathbb{Z}/\ell\mathbb{Z}) = \begin{cases}
\mathbb{Z}/\ell\mathbb{Z}, & \ell^\infty\nmid [L:K]\\
0, & \text{otherwise}
\end{cases} \]
\end{lemma}

\begin{proof}
Let $T$ be the absolute inertia group of $K$ and consider the short exact sequence from Theorem \ref{theorem-absoluteinertia}:
\[ 1\longrightarrow T\longrightarrow G_K\longrightarrow G_k\longrightarrow 1 \]

Let $G_\ell$ be an $\ell$-Sylow group of $G_K$. By Theorem \ref{theorem-sylowgroups-inertiagroup}, the $\ell$-Sylow group of $T$ isomorphic to $\mathbb{Z}_\ell^r$, where $r =\dim_{\mathbb{F}_\ell}(\Gamma/\ell\Gamma)$. Furthermore, $G_k\subset G_{\mathbb{F}_p}\cong\hat{\mathbb{Z}}$, thus the $\ell$-Sylow group $G_{k,\ell}$ of $G_k$ is either isomorphic to $\mathbb{Z}_\ell$ or trivial. The exact sequence above gives rise to the short exact sequence of their $\ell$-Sylow subgroups:
\[ 0\longrightarrow\mathbb{Z}_\ell^r \longrightarrow G_\ell \longrightarrow G_{k,\ell}\longrightarrow 1 \]
Since $G_{k,\ell}$ is a free pro-$\ell$ group, this sequence splits and we have an isomorphism $G_\ell \cong \mathbb{Z}_\ell^r\rtimes G_{k,\ell}$ with $r =\dim_{\mathbb{F}_\ell}(\Gamma/\ell\Gamma)\leq \operatorname{rrank}(\Gamma)\leq 2$.

>>>>

\documentclass[10pt,b5paper]{article}
\usepackage[left=2.00cm, right=2.00cm, top=2.00cm, bottom=2.00cm]{geometry}
\usepackage{fontspec}
\usepackage[ngerman]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{paralist}
\usepackage{tikz}
\usetikzlibrary{cd, babel}
\usepackage{hyperref}
\hypersetup{
	bookmarksopen=true,
	pdfstartview=FitH
}
\linespread{1.1}

\author{von \textsc{yichuan shen}}
\title{Galoissche Kennzeichnung von Funktionenkörper \\ vom Transzendenzgrad $1$ über Zahlkörper}
\begin{document}

\theoremstyle{plain}
\theoremstyle{definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Satz}
\newtheorem{corollary}[theorem]{Korollar}
\theoremstyle{definition}
\newtheorem*{definition}{Definition}
\newtheorem*{example}{Beispiel}
\theoremstyle{remark}
\newtheorem*{remark}{Bemerkung}
\newtheorem{step}{Schritt}

\maketitle

\section{Valuations}

@@

@@

We can endow $\operatorname{Spv}(K)$ with a natural topology by declaring following sets as a basis of open sets:
\[\{v\in\operatorname{Spv}(K)\mid v(x)\geq 0\text{ for all }x\in S\} \]
where $S$ ranges over all finite subsets of $K$. Let $v,w\in\operatorname{Spv}(K)$. We have:
\[\mathcal{O}_w\subset\mathcal{O}_v \iff w\in\overline{\{v\}} \]
Thus $v$ is a closed point if and only if it has no non-trivial specializations.
\end{definition}

@@

@@

\section{Main Lemma}

Let $K/\mathbb{Q}$ be a field of dimension $2$.

\begin{lemma}[\textit{Main Lemma}]\label{2.4}
Let $\ell\neq 2$ be a prime. The following map, induced by the restrictions, is injective:
\[ \mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z})\longrightarrow \prod_{w\in |\operatorname{Spv}(K)| }\mathrm{H}^3(K_w,\mathbb{Z}/\ell\mathbb{Z}) \]
\end{lemma}

\begin{lemma}\label{2.5}
Let $K$ be henselian with respect to $w\in|\operatorname{Spv}(K)|$. Set $k=k(w)$ and $\Gamma = \Gamma_w$. Then $k$ is an algebraic extension of a finite field $\mathbb{F}_p$. Let $\ell\neq p$ be a prime and assume $\mu_\ell\subset K$. Then:
\[ \mathrm{H}^3(K, \mathbb{Z}/\ell\mathbb{Z}) = \begin{cases}
\mathbb{Z}/\ell\mathbb{Z}, & \ell^\infty\nmid [k:\mathbb{F}_p] \text{ and } \dim_{\mathbb{F}_\ell}(\Gamma / \ell\Gamma) = 2\\
0, & \text{otherwise}
\end{cases} \]
If $L/K$ is an algebraic extension and $\mathrm{H}^3(K, \mathbb{Z}/\ell\mathbb{Z}) = \mathbb{Z}/\ell\mathbb{Z}$ holds, we have:
\[ \mathrm{H}^3(L,\mathbb{Z}/\ell\mathbb{Z}) = \begin{cases}
\mathbb{Z}/\ell\mathbb{Z}, & \ell^\infty\nmid [L:K]\\
0, & \text{otherwise}
\end{cases} \]
\end{lemma}
